---
title: "Data Reformatting practice"
author: "SK"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Two data tables with species and region definition are not very tidy as species table with individual species in the column so needs to move to have species column and count column.

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)

# to call stats filter - stats::filter()
```

[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_ori <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", stringsAsFactors = F)
region_ori <- read.csv("https://knb.ecoinformatics.org/view/df35b.304.2", stringsAsFactors = F)
```

# Reformat Catch Data

* remove "all" and "notes" columns.
* create "species" column.
    - move from wide to tall
* general QA

****
Remove the "all" column by using "select"" function and pipe operator (Cmd + Shift + m)

```{r}
catch_long <- catch_ori %>% 
  select(-All, -notesRegCode) %>% #select columns we want
  gather(key = Species, value = Catch, -Region, -Year) #move wide to long
#select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  
head(catch_long)
#select(.data, ...)
```

Move back to wide
```{r}
catch_wide <- catch_long %>% 
  spread(key = Species, value = Catch)
head(catch_wide)
```

## Clean up our data or general QA

* rename catch to catch_thousand
* change catch column to numeric
* create a new catch column in units num. of fish

```{r}
catch_clean <- catch_long %>% 
  rename(catch_thousands = Catch) %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  # %in% for multiple call or use correction table and join
  mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
  mutate(Catch = catch_thousands * 1000) %>% 
  select(-catch_thousands)

head(catch_clean)
```

```{r, eavl = F, echo = F}
#Look for where as.numeric failed
test <- as.numeric(catch_long$Catch)
i <- which(is.na(test) == T)
catch_long[i,] #letter I is confused with number 1 in GSE region for 1955 (401 case)
```

## Split - Apply - Combine

* calculate mean catch by species
```{r}
species_mean <- catch_clean %>% 
  group_by(Species) %>% 
  summarise(Catch_mean = mean(Catch),
            Num_obs = n())
#            Num_obs = length(Catch > 0))
#            Num_obs = length(which(is.na(Catch) == F))
head(species_mean)
```

* calculate mean catch by year
```{r}
species_mean <- catch_clean %>% 
  group_by(Year) %>% 
  summarise(Catch_mean = mean(Catch)) %>% # mutate adds
  arrange(-Catch_mean) %>% 
  filter(Year >= 1990 & Year <= 1995) #or is |

head(species_mean)
```

# Join the Region Definitions




# Misc. Function
