---
title: "Data Reformatting practice"
author: "SK"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Two data tables with species and region definition are not very tidy as species table with individual species in the column so needs to move to have species column and count column.

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)

# to call stats filter - stats::filter()
```

[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_ori <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", stringsAsFactors = F)
region_ori <- read.csv("https://knb.ecoinformatics.org/view/df35b.304.2", stringsAsFactors = F)
```

# Reformat Catch Data

* remove "all" and "notes" columns.
* create "species" column.
    - move from wide to tall
* general QA

****
Remove the "all" column by using "select"" function and pipe operator (Cmd + Shift + m)

```{r}
catch_long <- catch_ori %>% 
  select(-All, -notesRegCode) %>% #select columns we want
  gather(key = Species, value = Catch, -Region, -Year) #move wide to long
#select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  
head(catch_long)
#select(.data, ...)
```

Move back to wide
```{r}
catch_wide <- catch_long %>% 
  spread(key = Species, value = Catch)
head(catch_wide)
```


# Join the Region Definitions

# Misc. Function
