---
title: "Data Reformatting practice"
author: "SK"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Two data tables with species and region definition are not very tidy as species table with individual species in the column so needs to move to have species column and count column.

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)

# to call stats filter - stats::filter()
```

[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_ori <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", stringsAsFactors = F)
region_ori <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", stringsAsFactors = F)
#for windows system, region_defs<-read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),stringsAsFactors = FALSE)
```

# Reformat Catch Data

* remove "all" and "notes" columns.
* create "species" column.
    - move from wide to tall
* general QA

****
Remove the "all" column by using "select"" function and pipe operator (Cmd + Shift + m)

```{r}
catch_long <- catch_ori %>% 
  select(-All, -notesRegCode) %>% #select columns we want
  gather(key = Species, value = Catch, -Region, -Year) #move wide to long
#select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  
head(catch_long)
#select(.data, ...)
```

Move back to wide
```{r}
catch_wide <- catch_long %>% 
  spread(key = Species, value = Catch)
head(catch_wide)
```

## Clean up our data or general QA

* rename catch to catch_thousand
* change catch column to numeric
* create a new catch column in units num. of fish

```{r}
catch_clean <- catch_long %>% 
  rename(catch_thousands = Catch) %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  # %in% for multiple call or use correction table and join
  mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
  mutate(Catch = catch_thousands * 1000) %>% 
  select(-catch_thousands)

head(catch_clean)
```

```{r, eavl = F, echo = F}
#Look for where as.numeric failed
test <- as.numeric(catch_long$Catch)
i <- which(is.na(test) == T)
catch_long[i,] #letter I is confused with number 1 in GSE region for 1955 (401 case)
```

## Split - Apply - Combine

* calculate mean catch by species
```{r}
species_mean <- catch_clean %>% 
  group_by(Species) %>% 
  summarise(Catch_mean = mean(Catch),
            Num_obs = n())
#            Num_obs = length(Catch > 0))
#            Num_obs = length(which(is.na(Catch) == F))
head(species_mean)
```

* calculate mean catch by year
```{r}
species_mean <- catch_clean %>% 
  group_by(Year) %>% 
  summarise(Catch_mean = mean(Catch)) %>% # mutate adds
  arrange(-Catch_mean) %>% 
  filter(Year >= 1990 & Year <= 1995) #or is |

head(species_mean)
```

# Join the Region Definitions

Left join between catch and region definitions
```{r}
region_defs <- region_ori %>% 
  select(code, mgmtArea)

head (region_defs)
```

```{r}
Catch_joined <- left_join(catch_clean, region_defs, by = c("Region" = "code")) # or by = "Region" with rename (Region = code)

head (Catch_joined)
```
By comparing with inner_join tell you that there are missing info in right data.frame

# Misc. Function

Separate and Unite

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

```

```{r}
dates_split <-dates_df %>% 
  separate(date, into = c("Month", "Day", "Year"), sep = "/", remove = F)

head(dates_split)
```
This can be used to separate e.g., coded sample with multiple levels - TX_Wa_05, TX_Au-06...

```{r}
dates_unite <-dates_split %>% 
  unite(col = date_iso, Year, Month, Day, sep = "-")

head(dates_unite)
```
for date manipulation use lubridate package
stringr package is another one to consider for this kind of work - tidyverse

```{r}
library(stringr)

str_pad("5", 2, side = "left", pad = "0")
```

write data
```{r}
write.csv(catch_clean, "catch_clean.csv", row.names = F, quote = F)
```

****
mutate function can be used to create a column with row names for figure labeling purpose...