---
title: "Data Visualization"
author: "Sanghoon Kang"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reformat the data for plotting

[Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.](https://knb.ecoinformatics.org/view/doi:10.5063/F1S46Q6M)

```{r}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(scales)
```


```{r}
esc <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = F)

head(esc)
```

* calculate annual escapment by species and region
    + 'separate', 'group_by', 'summarize' using the 'sum' function
* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)

```{r}
annual_esc <-esc %>% 
  separate(sampleDate, into = c("Year", "Month", "Day"), sep = "-") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Year, Species, SASAP.Region) %>% 
  summarize(Count = sum(DailyCount)) %>% 
  filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum"))

head(annual_esc)

```

# Generate some static plots

```{r, echo = F}
ggplot(annual_esc, mapping = aes(x = Species, y = Count, fill = SASAP.Region)) +
#  scale_fill_brewer(palette = "Dark2") +
  geom_col() #data and mapping calls can be overridden here
```

```{r, echo = F}
my_theme <- theme_test() + #default theme, so put in front of other customization
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5, face="bold"))
```

```{r, echo = F}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), mapping = aes(x = Year, y = Count, color = Species)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = comma) +
  ylab("Escapement (num of fish)") +
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  theme_test() +
  my_theme
#  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r, fig.width = 7, fig.height = 10, , echo = F} 
ggplot(annual_esc, aes(Year, Count, color = Species)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = comma) +
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 2) +
  my_theme

#ggsave("figures/regional_escapement.png", 
#       plot = p,
#       width = 7,
#       height = 8,
#       units = "in")
```

# Generate an interative map

### Map the sampling locations for escapement

* find distinct lat/lon/location combinations

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na() #remove all the rows with NA

```

```{r, echo = F}
datatable(locations)
```

```{r}
leaflet(locations) %>% 
  addTiles() %>% # basemap title
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~Location)
```

'geom_sf()' for vector maps with sf package or ggmap
